---
title: "Reproducible research files"
output: 
  github_document:  # uncomment html lines and comment this line to render a README.html
  # html_document:
  #   toc: true
  #   toc_float: true
  #   toc_collapsed: true
  #   toc_depth: 3
  #   number_sections: true
  #   theme: lumen
  editor_options: 
    chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache=FALSE, 
  comment = "#>"
)
```

This README documents the procedures to replicate all analyses in the manuscript entitled 
"Regularized parametric survival modeling to improve risk prediction models".

Authors: J Hoogland, TPA Debray, MJ Crowther, RD Riley, J IntHout, JB Reitsma, AH Zwinderman
Code was written by J Hoogland

The required files are available as supplementary material on the Biometrical Journal website (excluding intermediate output files due to their size) and on GitHub (https://github.com/jeroenhoogland/regsurv-bimj-repro). To easily reproduce all figures, tables, and numerical results, please ensure that the intermediate files are downloaded and stored in folder `./intermediates/` (this is automatically the case when cloning the GitHub repository).

In case of questions or comments please contact j.hoogland@amsterdamumc.nl

# Software versions

The code was written/evaluated in R with the following software versions:

```{r, include=FALSE}
library(dplyr)
library(regsurv)
library(CVXR)
library(simsurv)
library(survival)
library(MASS)
library(glmnet)
library(rstpm2)
library(polspline)
library(Hmisc)
library(stringr)
```

```{r, echo=FALSE}
sessionInfo()
```

As an exception to the above, simulations were run in parallel on a Linux server with software versions:

```{r, echo=FALSE}
load("intermediates/survsim1.RData")
results$sessionInfo
```

# File structure

 * `README.html` the general README information you're now reading.
 * Folder `Results` contains 
   * `Results_bimj.Rmd` contains the Markdown script to generate `Results_bimj.html`, which generates all results for the manuscript.
 * Folder `./sim/` contains all R scripts pertaining to the simulation study
   * `dgm.R` contain the R script to generate all simulated data (files `groundtruth.RData`, `dev_population.RData`, and `validation.RData`) and the supplementary in the data generating mechanims section (`dgm.eps`).
   * `master_sim.R` Master file that runs all simulations and stores the results. Note that each run has its own random seed, such that individual runs can be replicated. NB: All sample size settings are generated from the same run.
   * `groundtruth.RData` contains the ground truth for the simulated data.
   * `dev_population.RData`contains the simulated development population (N=100,000)
   * `validation.RData` contains the simulated validation data
   * `helperFunctions.R` contains helper function that are not packaged (only useful to avoid code replication in the current context)
   * folder `./figures/` contain all manuscript figures in the required formats
   <!-- * Folder `./HPC_bimj/` contains scripts for the parallel runs on a linux server (not required) -->
     <!-- * `*.sh` shell scripts that contains the code to run the simulations on a CentOS High Performance Grid (HPC) using Slurm. -->
     <!-- * `*.R` contains the R scripts for a single run for each group of settings (main, high censoring, high p, applied example). -->
 * Folder `./ae/` contains all R scripts pertaining to the applied example
   * `ae_apparent.R` contains the R script for the main analyses
   * `ae_bimj_boot.R` contains the R script for a single bootstrap run
   * `helperFunctions_ae.R` contains helper function that are not packaged (only useful to avoid code replication in the current context).
   * `ae_apparent_bimj.RData` contains the pre-computed results of running `ae_apparent.RData`.
 * Folder `./functions` contains
   * The documented function `calibrate()` stored in `calibrate.R`, which can be loaded using `source("functions/calibrate.R")`
   * (Note that all analysis functions are directly loaded from R packages)
 * Folder `./intermediates/` contains all intermediate files (output files) from the high 
 performance grid for both the simulations (`survsim*.RData`, `survsim_highp*.RData`, 
 `survsim_highcens*.RData`) and the applied example (`ae*.RData`), where `*` is replaced by the seed number for the random number generator.
 
 
# Reproduce all figures and numerical results

The required computation time for the simulations is long (on average nearly 5 hours for the main simulation runs on the CentOS system and longer for the additional higher dimensional and increased censoring settings). 

Therefore, intermediate files are available for all simulation runs and the bootstrap runs for the applied example.

## The Results_bimj markdown and html

`Results_bimj.Rmd` contains all code to constructs all figures and numerical results from the intermediate files, and the corresponding `html` file `Results_bimj.html` provides the rendered output. Note that numerical results are provided in `Results_bimj.html` (such as Table 1), and that figures are not rendered but stored in the correct format in folder `./figures`. Each of the figures has the exact same file name as in the Latex files, such that they can be inserted in the right place without any modification. 

## Replicating the simulation data

R script `dgm.R` in folder `./sim/` generates all required simulated data and stored it in the right place. Also, it generates `dgm.eps`, which is the supplementary figure on the data generating mechanism. 

Resulting files are stored in folder `./data/` and are named `dev_population.RData` (the part of the simulated population from which development data sets / training data are sampled), `validation.RData` (the part of the simulated population reserved for validation purposes), and `groundtruth.RData` (which contains all necessary information on the true model and true survival probabilities).

## Replicating the intermediate files

First of all, R package **regsurv** needs to be installed from GitHub

```
library(devtools)
install_github("jeroenhoogland/regsurv")
```

### Simulations

In folder `./sim/`, R script `master_sim.R` can run all simulations (for all simulation settings). The specifications for the different simulation settings can be set within the master file and have intuitive names: `sample.size.settings` takes number 1:4 for each of the 4 sample size settings in the manuscript (N=100, 250, 500, 1000), `highercens` is TRUE for the high censoring settings, `higherp` is true for the setting with a larger set of covariates, and both are FALSE for the default simulation settings.

An impression of the script to run the main simulation settings is

```
# Master script to run the simulation study
nsim <- 1000

# Main simulations (NOT RUN)
for(iter in 1:nsim){
  set.seed(iter)
  
  sample.size.settings <- 1:4 # n=100, 250, 500, 1000
  highercens <- FALSE
  higherp <- FALSE 
  
  source("sim/sims.R")
  
  save(results, file= paste0("intermediates/", "survsim", iter, ".RData"))
}
```

R script `master_sim.R` also includes a few lines of code that show spot replication of the 10th simulation run for sample size N=100 on a different machine.

As a generic example, to replicate let's say run 98 for sample size setting 1 of the main simulations, the following code can be used:

```
iter <- 98
set.seed(iter)
sample.size.settings <- 1
highercens <- FALSE
higherp <- FALSE
source("sim/sims.R")
spot_check <- results
```

On a recent MBP (2021, Apple M1 Pro, 16GB working memory), this takes approximately 48 minutes.

This generates a results object that can be checked against `./intermediates/survsim98.RData`.

For instance, to check the overall root mean squared prediction errors for each of the methods, compare

```
load("intermediates/survsim98.RData") # loads the results object for sim 98
all.equal(spot_check$rmse[1,,], results$rmse[1,,], tolerance=10^-5) # should return TRUE
all.equal(sapply(spot_check$cstats[[1]], function(x) x[1, ]), 
          sapply(results$cstats[[1]], function(x) x[1, ]), tolerance=10^-5) # should return TRUE
```

### Applied example

The intermediate files for the applied example can also be reproduced. The apparent (whole sample) results can be replicated using R script `ae_apparent.R` in the `./ae/` folder. The bootstrap runs can be replicated using `ae_bimj_boot.R` in the same folder.

NB boot needs b and seed

For instance:

```
b <- 9 # for replication of bootstrap 9
set.seed(b)
source("ae/ae_bimj_boot.R")
spot_check <- boot
```

Checks against

```
load("intermediates/ae9.RData") 
unlist(spot_check$cstats.oob) # OOB Ctd estimates
all.equal(unlist(spot_check$cstats.oob), unlist(boot$cstats.oob), tol=10^-5)
```

END